# The `PSTricks` object

After installation, a first test could be, within R:
```{r}
names(PSTricks::PSTricks())
```

`PSTricks()` returns an (S3) object with various attributes. The
"lines" attribute will contain the LaTeX code lines. The attributes
are manipulated by the various PSTricks functions and normally need
not be accessed directly.  See the documentation of the `PSTricks()`
function for a short description of the shown attributes of the
`PSTricks` object.

Next, to see if running LaTeX works:
```{r eval=FALSE}
PSTricks::PSTricks()
```
However, because the document is empty, no .pdf will be generated.

To be able to view the generated "pp.tex" file:
```{r}
print(PSTricks(),"test",topdf=FALSE)
cat(system("cat test.tex",intern=TRUE),sep='\n')
```

A test that should give a .pdf (called "pp.pdf" if run in an
R session rather than a script):
```{r}
geom_dots(PSTricks(),aes(x=wt,y=mpg),mtcars)
```
```{r,echo=FALSE,fig.align="center"}
knitr::include_graphics("renderbook.pdf")
```

\newpage
The output is one or more whole pages by default, that is,
a picture is not "cropped" (see Chapter \@ref(printing)).

## Flow of `PSTricks` objects

- Creating an object:

```{r eval=FALSE}
p <- PSTricks()
```

See the main documentation for a description of the arguments
of `PSTricks()`.

- Modifying attributes:

```{r eval=FALSE}
p <- function(p, ...)
```

- Modifying using pipes (automatically imported from the "magritrr" package):

```{r eval=FALSE}
p <- p %>%
    functionA(...) %>%
    functionB(...)
```

- Printing the assembled information:

```{r eval=FALSE}
print(p)
```

See Chapter \@ref(printing) for the latter.

## Messages

Errors with piping the PSTricks object may cause confusing error messages;
for example, with

```{r eval=FALSE}
p <- p %>%
    functionA(...) %>%
    functionB(...) %>%

print(p)
```
`p` will be the first and second argument of `ppwrite()`, and the latter function
expects a string as the second argument.

See also directory "examples/errors/".

## Returning a string from a wrapped PSTricks macro

Most functions add a line to the "lines" attribute of the PSTricks object,
and to access that attribute, the object variable ("p" in the above) has
to be available.

Some PSTricks macro calls are nested, so that sometimes the macro call
as a string is needed and that string should not be appended to the lines
attribute. To allow this, the object may be specified as NULL, which
is the default if possible, so for example:

```{r}
psline(,c(1,2),c(3,4))
```

See Chapter \@ref(postscript) for a complex example.
